---
title: "Reproducible Research : Assignment 1"
output:
  html_document:
    keep_md: true
---

#### Environment Details

R version 3.1.0 (2014-04-10)
Platform: x86_64-w64-mingw32/x64 (64-bit)

locale:
[1] LC_COLLATE=English_Australia.1252  LC_CTYPE=English_Australia.1252   
[3] LC_MONETARY=English_Australia.1252 LC_NUMERIC=C                          
[5] LC_TIME=English_Australia.1252       

attached base packages:  
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:  
[1] boot_1.3-11

loaded via a namespace (and not attached):  
[1] tools_3.1.0


Reading the activity file and converting Date column from factor to Date format
```{r, echo =TRUE}
activity <- read.csv("./data/activity.csv")
activity[,2]<-as.Date(activity[,2],format="%Y-%m-%d")

```

Plotting the total steps per day
```{r, echo=TRUE}
a<-aggregate( steps~date, activity, sum, na.action=na.pass )
hist(a$steps, breaks=25, xlab="Total Steps", main="Summary of steps", col="blue")
mean_of_daily_totals <- as.character(mean(a[,2],na.rm=TRUE))
median_of_daily_totals <- as.character(median(a[,2],na.rm=TRUE))

```
The mean is `r mean_of_daily_totals`  
The median is `r median_of_daily_totals`


Average daily activity pattern  

```{r, echo=TRUE}
b<-aggregate( steps~interval, activity, mean, na.action=na.omit )
plot(b$interval, b$steps, type="l", col="blue", xlab="Interval", ylab="Average steps", main="")
b1<- subset(b,b$steps==max(b$steps))
print(b1$interval)
```
The interval with the maximum number of steps is `r b1$interval`

#### Imputing missing values  
Derive values to impute based on average per interval over the entire dataset    
Create the new dataset with the imputed values in place of "NA"  
Make a histogram with total steps taken each day    
compare with first step the difference in mean and median  

```{r, echo=TRUE}
##c<-subset(activity, is.na(steps) | is.na(date) | is.na(interval))
c1<-merge(activity, b , by="interval")
for(i in 1:length(c1$steps.x))
    {if (is.na(c1$steps.x[i])) 
        {c1$steps.x[i] <- c1$steps.y[i]
        }
    }
c1 <- subset(c1, select=c(1:3))
c1<-aggregate( steps.x~date, c1, sum)
hist(c1$steps.x, breaks=25, xlab="Total Steps", main="Summary of imputed steps", col="blue")
print(as.character(mean(c1[,2])))
print(as.character(median(c1[,2])))
```
