---
title: "Reproducible Research : Assignment 1"
output:
  html_document:
    keep_md: true
---

#### Environment Details

```{r, echo =TRUE}
sessionInfo()
```

#### Start of the assignment 

#####  Loading and preprocessing the data  

Reading the activity file and converting Date column from factor to Date format
```{r, echo =TRUE}
activity <- read.csv("./data/activity.csv")
activity[,2]<-as.Date(activity[,2],format="%Y-%m-%d")

```
  
##### Plotting the total steps per day  
```{r, echo=TRUE}
a<-aggregate( steps~date, activity, sum, na.action=na.pass )
hist(a$steps, breaks=25, xlab="Total Steps", main="Summary of steps", col="blue")
mean_of_daily_totals <- as.character(round((mean(a[,2],na.rm=TRUE)), digits=2))
median_of_daily_totals <- as.character(round((median(a[,2],na.rm=TRUE)), digits=2))

```
The mean is `r mean_of_daily_totals`  
The median is `r median_of_daily_totals`


##### Average daily activity pattern    

```{r, echo=TRUE}
b<-aggregate( steps~interval, activity, mean, na.action=na.omit )
plot(b$interval, b$steps, type="l", col="blue", xlab="Interval", ylab="Average steps", main="")
b1<- subset(b,b$steps==max(b$steps))
print(b1$interval)
```
The interval with the maximum number of steps is `r b1$interval`

##### Imputing missing values    
Derive values to impute based on average per interval over the entire dataset    
Create the new dataset with the imputed values in place of "NA"  
Make a histogram with total steps taken each day    
compare with first step the difference in mean and median  

```{r, echo=TRUE}
##c<-subset(activity, is.na(steps) | is.na(date) | is.na(interval))
c1<-merge(activity, b , by="interval")
for(i in 1:length(c1$steps.x))
    {if (is.na(c1$steps.x[i])) 
        {c1$steps.x[i] <- c1$steps.y[i]
        }
    }
c1 <- subset(c1, select=c(1:3))
c2<-aggregate( steps.x~date, c1, sum)
hist(c2$steps.x, breaks=25, xlab="Total Steps", main="Summary of imputed steps", col="blue")
impMean<- as.character(round(mean(c2[,2]),digits=2))
impMedian<- as.character(round(median(c2[,2]),digits=2))
```
The mean of the orginal dataset was '`r mean_of_daily_totals`' and with the imputed values it is '`r impMean`'  
The median of the orginal dataset was '`r median_of_daily_totals`' and with  the imputed values it is '`r impMedian`'  

##### Plotting differences in activity patterns between weekdays and weekends

```{r, echo=TRUE}

d<- transform(c1, weekend=as.POSIXlt(date, format='%Y/%m/%d')$wday %in% c(0, 6))
wEnd<- subset(d,d$weekend==TRUE)
wDay<- subset(d,d$weekend==FALSE)

wEndAve<-aggregate( steps.x~interval+date, wEnd, mean)
wDayAve<-aggregate( steps.x~interval+date, wDay, mean)

par(mfrow=c(2,1))
plot(wEndAve$interval, wEndAve$steps.x, type="l", col="blue", main="Weekend", xlab="Interval", ylab="Number of steps")
plot(wDayAve$interval, wDayAve$steps.x, type="l", col="blue", main="Weekday", xlab="Interval", ylab="Number of steps")

```
